\hypertarget{design-by-contract}{%
\section{Design by Contract}\label{design-by-contract}}

\begin{itemize}
\tightlist
\item
  A contract defines obligations of both parties as well as their
  benefits
\item
  It also defines what happens if obligations aren't fulfilled.
\item
  In software both parties are components

  \begin{itemize}
  \tightlist
  \item
    class, package, function, method
  \end{itemize}
\end{itemize}

\hypertarget{structure}{%
\subsection{Structure}\label{structure}}

\begin{itemize}
\tightlist
\item
  Preconditions

  \begin{itemize}
  \tightlist
  \item
    what the client needs to provide as true
  \item
    require from the client
  \end{itemize}
\item
  Postconditions

  \begin{itemize}
  \tightlist
  \item
    what the component promises to establish
  \item
    ensure to the client
  \end{itemize}
\item
  Invariants

  \begin{itemize}
  \tightlist
  \item
    are conditions that remain true for the component implementing the
    methods, regardless what happens
  \end{itemize}
\end{itemize}

\hypertarget{preconditions}{%
\subsection{Preconditions}\label{preconditions}}

\begin{itemize}
\tightlist
\item
  must be true before call is made
\item
  client (caller) is responsible
\item
  supplier mustn't re-check preconditions
\item
  Opposite of ``Defensive Programming''
\end{itemize}

\hypertarget{postconditions}{%
\subsection{Postconditions}\label{postconditions}}

\begin{itemize}
\tightlist
\item
  defines condition that is true after a call is finished
\item
  responsibility lies in the hand of the supplier (method)
\item
  when an exception can be thrown, there are two kinds of postconditions

  \begin{itemize}
  \tightlist
  \item
    normal postconditions

    \begin{itemize}
    \tightlist
    \item
      define behavior with regular control flow
    \end{itemize}
  \item
    exceptional postconditions

    \begin{itemize}
    \tightlist
    \item
      define behavior when an exception was thrown
    \end{itemize}
  \end{itemize}
\end{itemize}

\hypertarget{contracts-and-inheritance}{%
\subsection{Contracts and Inheritance}\label{contracts-and-inheritance}}

\begin{itemize}
\tightlist
\item
  Contracts belong to the interface
\item
  implementing classes must fulfill given contract
\end{itemize}

\hypertarget{six-principles-for-dbc}{%
\subsection{Six Principles for DbC}\label{six-principles-for-dbc}}

\begin{itemize}
\tightlist
\item
  Separate queries from commands.

  \begin{itemize}
  \tightlist
  \item
    Queries return a result but do not change the visible properties of
    the object.
  \item
    Commands might change the object but do not return a result.
  \end{itemize}
\item
  Separate basic queries from derived queries.

  \begin{itemize}
  \tightlist
  \item
    Derived queries can be specified in terms of basic queries.
  \end{itemize}
\item
  For each derived query, write a postcondition that specifies what
  result will be returned in terms of one or more basic queries.

  \begin{itemize}
  \tightlist
  \item
    Then, if we know the values of the basic queries, we also know the
    values of the derived queries.
  \end{itemize}
\item
  For each command, write a postcondition that specifies the value of
  every basic query.

  \begin{itemize}
  \tightlist
  \item
    Now we know the total visible effect of each command.
  \end{itemize}
\item
  For every query and command decide on a suitable precondition.

  \begin{itemize}
  \tightlist
  \item
    Preconditions constrain when clients may call the queries and
    commands.
  \end{itemize}
\item
  Write invariants to define unchanging properties of objects.

  \begin{itemize}
  \tightlist
  \item
    Concentrate on properties that help the reader build an appropriate
    conceptual model of the abstraction the class embodies.
  \end{itemize}
\end{itemize}

\hypertarget{problems-with-dbc}{%
\subsection{Problems with DbC}\label{problems-with-dbc}}

\begin{itemize}
\tightlist
\item
  Lack of language support (except for Eiffel)
\item
  Not used systematically
\end{itemize}

\hypertarget{benefits-of-dbc}{%
\subsection{Benefits of DbC}\label{benefits-of-dbc}}

\begin{itemize}
\tightlist
\item
  Tool for specifying software components in a sound way

  \begin{itemize}
  \tightlist
  \item
    contracts abstract from implementation
  \item
    good documentation
  \end{itemize}
\item
  Clearly defined responsibilities between client and supplier
\item
  Simpler code that would be required from defensive programming
\end{itemize}

\clearpage