\hypertarget{introduction-to-agile-and-extreme-programming}{%
\section{Introduction to agile and Extreme
Programming}\label{introduction-to-agile-and-extreme-programming}}

Today a lot of software projects fail. This is because of the used
(heavyweighted) methologies, project management and so on. The evolution
of technology (hardware) and tools (IDEs etc.) has outstripped
methodologies by an order of magnitude.

\hypertarget{the-agile-manifesto}{%
\subsection{The Agile Manifesto}\label{the-agile-manifesto}}

The Agile Manifesto was written by some of the most famous developers.

\begin{itemize}
\tightlist
\item
  We are uncovering better ways of developing software by doing it and
  helping others do it
\item
  Our highest priority is to satisfy the customer through early and
  continuous delivery of valuable software.
\item
  Welcome changing requirements, even late in development.
\item
  Deliver working software frequently
\item
  Business people and developers must work together daily throughout the
  project
\end{itemize}

\subsubsection{Principles}
\begin{itemize}
    \item In Intervall the team reflects the work to become more effective.
    \item Working software is the unit of measurement
    \item Most effective method to receive information is face to face communication.
    \item Use it, measure it, remove it if its not used.
    \item The best architectures, requirements, and designs emerge from self-organizing teams.
    \item Simplicity - the art of maximizing the amount of work not done - is essential and elegance.
    \item Continuous attention to technical excellence and good design enhances agility.
    \item Build projects around motivated individuals. Give them the environment, support and trust they need.
\item Business and Developers must work together
\item Deliver on shorter timescale (couple of weeks or months)
\item Highest Priority is to satisfy the customer with continuous delivery
\end{itemize}















\hypertarget{extreme-programming-xp}{%
\subsection{eXtreme Programming (XP)}\label{extreme-programming-xp}}

\hypertarget{the-problem-risks-in-development-projects}{%
\subsubsection{The problem: risks in development
projects}\label{the-problem-risks-in-development-projects}}

\begin{itemize}
\tightlist
\item
  Schedule slips
\item
  Project canceled
\item
  System does not evolve gracefully
\item
  Business misunderstood
\item
  Business changes
\end{itemize}

4 variables (and their relations) while developing a software:

\begin{itemize}
\tightlist
\item
  Time
\item
  Resources
\item
  Quality
\item
  Scope
\end{itemize}

But the problem is, often 3 of these 4 variables are defined and part of
the contract (time, resources, scope). And if your project is running
late, you have to cut edges somewhere, so it's the quality.

\hypertarget{the-values-of-xp}{%
\subsubsection{The values of XP}\label{the-values-of-xp}}

\hypertarget{simplicity}{%
\paragraph{Simplicity}\label{simplicity}}

\begin{itemize}
\tightlist
\item
  We will do what is needed and asked for, but no more.
\item
  This will maximize the value created for the investment made to date.
\item
  We will take small simple steps to our goal and mitigate failures as
  they happen.
\item
  We will create something we are proud of and maintain it long term for
  reasonable costs.
\end{itemize}

\hypertarget{communication}{%
\paragraph{Communication}\label{communication}}

\begin{itemize}
\tightlist
\item
  Everyone is part of the team and we communicate face to face daily.
\item
  We will work together on everything from requirements to code.
\item
  We will create the best solution to our problem that we can together.
\end{itemize}

\hypertarget{feedback}{%
\paragraph{Feedback}\label{feedback}}

\begin{itemize}
\tightlist
\item
  We will take every iteration commitment seriously by delivering
  working software.
\item
  We demonstrate our software early and often then listen carefully and
  make any changes needed.
\item
  We will talk about the project and adapt our process to it, not the
  other way around.
\end{itemize}

\hypertarget{courage}{%
\paragraph{Courage}\label{courage}}

\begin{itemize}
\tightlist
\item
  We will tell the truth about progress and estimates.
\item
  We don't document excuses for failure because we plan to succeed.
\item
  We don't fear anything because no one ever works alone.
\item
  We will adapt to changes when ever they happen
\end{itemize}

\hypertarget{respect}{%
\paragraph{Respect}\label{respect}}

\begin{itemize}
\tightlist
\item
  Everyone gives and feels the respect they deserve as a valued team
  member.
\item
  Everyone contributes value even if it's simply enthusiasm.
\item
  Developers respect the expertise of the customers and vice versa.
\item
  Management respects our right to accept responsibility and receive
  authority over our own work.
\end{itemize}

\hypertarget{xp-practices}{%
\subsubsection{XP Practices}\label{xp-practices}}

\begin{itemize}
\tightlist
\item
  The Planning Game (estimate the effort)
\item
  Pair programming (Two people looking at the same screen)
\item
  Small releases (Every release should be as small as possible)
\item
  Collective ownership (leave the code better than you found it)
\item
  Metaphor (Common understanding of the system)
\item
  Continuous integration (Code is integrated and tested)
\item
  Simple design (Put in what need when you need it)
\item
  40 hours week (suistainable development)
\item
  Testing (Tests are required for a later refactoring)
\item
  On-site customer (A real customer must be physically with the team)
\item
  Refactoring (improve the existing source code)
\item
  Coding standards (coding practices must be unified in the team)
\end{itemize}


\hypertarget{Differences XP <-> Scrum}{%
\subsection{Differences XP <-> Scrum}\label{differnces_xp_scrum}}
\begin{itemize}
  \item 
  Scrum: Iterations called Sprints 2-4 Weeks, XP: 1-2 Weeks
  \item
    \textbf{Scrum teams}: No changes allowed into their sprint. Fixed at planning meetings, maybe do more.
    \textbf{XP teams}: Change within their iterations. As long as the team hasn’t started work on a particular feature.
  \item
    \textbf{Scrum}: Strict priority order in iteratio, prioritized by Product owner
    \textbf{XP}: No fixed priority order inside iteration
  \item
    \textbf{Scrum}: doesn’t prescribe any engineering practices; 
    \textbf{XP}: does
\end{itemize}